---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/readme/README-"
)
```

[![Build Status](https://travis-ci.org/GLEON/rLakeAnalyzer.svg?branch=master)](https://travis-ci.org/GLEON/rLakeAnalyzer)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/rLakeAnalyzer)](https://cran.r-project.org/package=rLakeAnalyzer) 
[![Download Count](http://cranlogs.r-pkg.org/badges/rLakeAnalyzer)]( https://cran.r-project.org/web/packages/rLakeAnalyzer/index.html)
[![DOI](https://zenodo.org/badge/23404/GLEON/rLakeAnalyzer.svg)](https://zenodo.org/badge/latestdoi/23404/GLEON/rLakeAnalyzer)

# rLakeAnalyzer

The `rLakeAnalyzer` package is an R implementation of the Lake Analyzer, a tool that allows users to calculate common metrics for lake physical states.

## Installation

You can install a stable version of rLakeAnalyzer with:
```{r cran-install, eval = FALSE}
install.packages("rLakeAnalyzer")
```

You can install the development version rLakeAnalyzer from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("boshek/rLakeAnalyzer")
```

## Usage
First the package must be loaded into R:
```{r load_in}
library(rLakeAnalyzer)
```

You can view a list of functions available in rLakeAnalyzer with:
```{r, eval = FALSE}
help(package = "rLakeAnalyzer")
```

An example application of this functionality is illustrated with `wtr.heat.map()` function:

```{r example}
#Get the path for the package example file included
wtr.path <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng Lake, Wisconsin.
sp.wtr = load.ts(wtr.path)

#Plot default figure
wtr.heat.map(sp.wtr)

#Change defaults supplied to filled.contour
wtr.heat.map(sp.wtr, zlim=c(0,15), plot.title="Sparkling Water Temp (C)")
```

## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/GLEON/rLakeAnalyzer/issues/).

These are very welcome!

## Contributors 
Luke Winslow, Richard Woolway, Jordan Read, Jennie Brentrup, Jake Zwart, Craig Snortheim, Taylor Leach, Sam Albers, Doug Collinge