---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
[![Travis-CI Build Status](https://travis-ci.org/boshek/limnotools.svg?branch=sam_exp)](https://travis-ci.org/boshek/limnotools) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/boshek/limnotools?branch=sam_exp&svg=true)](https://ci.appveyor.com/project/boshek/limnotools)


## limnotools package 
limnotools is an R package that implements the split and merge algorithm to high resolution water column profile to identify water column structure, stratification and mixing patterns. 

See: [limnotools vignette](https://github.com/boshek/limnotools/blob/master/vignettes/limnotools.md)

You can install the development version of limnotools from github with:
```{r, eval=FALSE}
devtools::install_github("boshek/limnotools")
```

## Important references
- Pavlidis, T., and S. L. Horowitz, 1974: Segmentation of plan curves.IEEE Trans. Comput., C-23, 860–870.
- Thomson, R. and I. Fine. 2003. Estimating Mixed Layer Depth from Oceanic Profile Data. Journal of Atmospheric and Oceanic Technology. 20(2), 319-329.
- Fiedler, Paul C. "Comparison of objective descriptions of the thermocline. Limnology and Oceanography: Methods 8.6 (2010): 313-325.

## Releases
- 0.9.0 Initial release
- 0.9.1 Fixed a fatal bug in wtr_layer and wtr_segment that fed incorrect measurement values in the call to by_s_m and by_s_m3
