---
title: "Limnotools Usage"
author: "Sam Albers and Doug Collinge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Packages needing loading
```{r, message=FALSE}
library(limnotools)
library(knitr)
```

### Development version of rLakeAnalyzer
At this time the split and merge algorithm is only available with the development version of `rLakeAnalyzer`. To install `rLakeAnalyzer`, you need the devtools package installed. Then you can install `rLakeAnalyzer` in R from github like this:
```{r, message=FALSE, eval=FALSE}
devtools::install_github("GLEON/rLakeAnalyzer", ref="devel")
library(rLakeAnalyzer)
```

## Split and merge algorithm
Water column identification is provided by the split-and-merge algorithm. Implementation of the split-and-merge algorithm for a water profile occurs within the `water_layer()` function:

## Simple application of the split and merge algorithm
Below is a simple one profile example of determining key water column parameters using the split-and-merge algorithm. The default behaviour for `water_layer` is to run the algorithm *without* specifying the number of segments. `water_layer()` adopt as defaults the convention of a minimum depth (z0) of 2.5 m, a maximum depth (zmax) of 150 m and a error threshold (thres) of 0.1. 
```{r}
wldf <- wtr_layer(depth = latesummer$depth, measure = latesummer$temper)
knitr::kable(wldf)
```

Note that the axes of the water column profile have been reversed and flipped to better visualize the water column and conform to standard limnological displays. 
```{r, fig.show = "hold", fig.width = 8, fig.height = 6}
plot(y = latesummer$depth, x = latesummer$temper, ylim = rev(range(latesummer$depth)))
abline(h = wldf$cline, col='blue')
abline(h = wldf$mld, col='red')
abline(h = wldf$min_depth, col='green')
text(16, wldf$cline+3, "Thermocline", col = 'blue')
text(16, wldf$mld+3, "Mix Layer Depth", col = 'red')
text(16, wldf$min_depth+3, "Minimum Depth", col = 'green')
```

